FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-08-15 paper.tex
Many big points addressed

@@

\FT{} has two roles:

\begin{enumerate}
  \item Track exceptional values
  \item Fuzz \fp{} code from the inside-out
\end{enumerate}

Building the mechanism to track exceptional values also gave us a way to inject exceptional values wherever \fp{} operations occur.
We will examine each of these roles, how they are used, and how \FT{} realizes them.

%% What it does, then how it works

@@

% note: there's a section on performance in the "discussion" section at the end, so we can leave it out here I think

While other tools like profilers might accomplish this tracking by writing extensions to the language's compiler or runtime, we were able to implement \FT{} entirely in user land.
We created custom types \texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64} that wrap their corresponding \texttt{Float16}, \texttt{Float32}, and \texttt{Float64} counterparts.
We then defined methods for every function in the \texttt{Base} module, including new implementations for operators like \texttt{+} and \texttt{*}.
This lets us intercept all \fp{} operations working on our \texttt{TrackedFloat} types.

@@

@@

CSTG~\cite{humphreySystematicDebuggingMethods2014} is built on the work of and related to the STAT~\cite{arnoldStackTraceAnalysis2007} tool developed by Laurence Livermore National Laboratory.
Please see the CSTG paper for more on the differences between the two.

@@

\subsection{Tracking more than \NaN{}s and \Inf{}s}
FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to \NaN{} or \Inf{}.
We spent most of the time engineering and testing \NaN{} tracking and injection.
Adding \Inf{} tracking took a few hours.
Adding other values of interest presents no fundamental difficulties.

@@

\subsection{Static FPX, Gradual Types}

FILL

@@

We would also like to thank the developers of \code{RxInfer.jl} for feedback on \FlowFPX{}.

@@

% Below is some Emacs-specific stuff. If you're using Emacs, you should try
% using the new jinx package for spell-checking!

>>>>

%% TODO items
%%  - [ ] Normalize casing in all sections, sub-sections, etc

@@

\FT{} is implemented as a Julia library, and is available for download on JuliaHub.\footnote{\url{https://juliahub.com/ui/Packages/FloatTracker/dBXig/1.0.0}}

@@

@@

\subsubsection{\FT{} internals}

While other tools like profilers might accomplish the tracking that \FT{} does by writing extensions to the language's compiler or runtime, we were able to implement \FT{} entirely in user land.
We created custom types \texttt{TrackedFloat16}, \texttt{TrackedFloat32}, and \texttt{TrackedFloat64} that wrap their corresponding \texttt{Float16}, \texttt{Float32}, and \texttt{Float64} counterparts.
We then defined methods for every function in the \texttt{Base} module, including new implementations for operators like \texttt{+} and \texttt{*}.
This lets us intercept all \fp{} operations working on our \texttt{TrackedFloat} types.

@@

CSTG~\cite{humphreySystematicDebuggingMethods2014} is built on the work of and related to the STAT~\cite{arnoldStackTraceAnalysis2007} tool developed by Laurence Livermore National Laboratory.
STAT is a tool for collecting, analyzing, and visualizing stack traces from thousands of concurrent processes, and is designed to highlight anomalies quickly.
It produces visualizations similar to those of CSTG, thought CSTG offers more compact views.

@@

\subsection{Tracking more than \NaN{}s and \Inf{}s}

FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to \NaN{} or \Inf{}.
We spent most of the time engineering and testing \NaN{} tracking and injection.
Adding \Inf{} tracking took a few hours.
Adding other values of interest presents no fundamental difficulties.

@@

\subsection{Tracking exceptions in external libraries}

\FT{}'s scope is limited to Julia code: it cannot track the lifetime of exceptional values that are born, live, or die in external libraries, such as C programs.
One approach would be to create an API for \FT{} so that programs using these external libraries could leverage \FT{}'s tooling to monitor the interface between Julia and e.g. external C programs.

@@

We would like to thank Alex Larsen and Rob Durst for their help with reading drafts of the paper and feedback on the presentation.
We would also like to thank the developers of \code{RxInfer.jl} for feedback on \FlowFPX{}.
Finally, many thanks to Milan Klöwer, the author of the \texttt{Sherlogs} library, for inspiration on the architecture of \FT{}.

@@

% Below is some Emacs-specific stuff to help me (Ashton) write this paper. If
% you're using Emacs, you should try using the new jinx package for
% spell-checking!
